Return on investment of private sector malaria control at a large sugar facility in Southern Mozambique

### 2018-01-04 inst/rmd/paper.tex
formatting updates

\usepackage[table]{xcolor}

% Font
\usepackage{fontspec}
\setmainfont{Swift-Regular_43151.ttf}
\setsansfont{Swift-Regular_43151.ttf}
% \setsansfont[BoldFont={Swift-Bold_43130.ttf}]{Swift-Bold_43130.ttf}
\setmonofont{Swift-Regular_43151.ttf}
\renewcommand{\familydefault}{\sfdefault}

\usepackage{multicol}
\usepackage{hyperref}
\usepackage{geometry}
\usepackage{lipsum}
\usepackage[utf8]{inputenc}

% \usepackage{todonotes} % for side notes
\usepackage[colorinlistoftodos]{todonotes} % for side notes
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{fancyhdr} % Headers and footers
\pagestyle{fancy} % All pages have headers and footers
\fancyhead{} % Blank out the default header
\fancyfoot{} % Blank out the default footer
\fancyhead[C]{Effectiveness of private sector malaria control at a large sugar
  facility in Southern Mozambique}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}

@@

\DeclareTextCommandDefault{\nobreakspace}{\leavevmode\nobreak\ }

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage[table]{xcolor}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}

@@

Malaria has a large economic impact on endemic societies. By affecting
saving, investment (Shretta, Avanceña, and Hatefi 2016), risk
perception, productivity, absenteeism (Nonvignon et al. 2016), human
capital accumulation (Castel-Branco 2014), mortality, and costs of care
(Sachs and Malaney 2002), malaria likely has a negative effect on GDP
and growth (McCarthy, Wolf, and Wu 2000) (Orem et al. 2012). Because of
the relative affordability of most intereventions and the enormous
societal costs of malaria, most forms of malaria control are
cost-effective when a public welfare perspective is assumed, such as
when a government provides the financing (White et al. 2011) (Purdy et
al. 2013) (Howard et al. 2017).

From the persepctive of the private sector, however, investing in
malaria control is not so clear-cut. Public health interventions
targetting malaria - and their corresponding cost-effectiveness
evaluations - most often focus on impacts pertaining to public welfare,
such as an increase in life years adjusted for disability or quality
(Goodman, Coleman, and Mills 1999) (Shretta, Avanceña, and Hatefi 2016)
(Lee et al. 2017) (Hanson 2004). Though population-level health is
certainly of importance to businesses, and improvements in health
incidentally improve the economy at all levels (Brundtland 1999) (Bloom
and Canning 2008) (Vecchi, Hellowell, and Gatti 2013), these
improvements may be too disperse or long-term to incentivize private
sector involvement in health campaigns.

100\% of the Mozambican population are at risk of malaria, living in
what the WHO classifies as a ``high transmission'' area (Moonasar et al.
2016). Annually, Mozambique has more than 8 million clinical malaria
cases (an annual incidence of approximately 300 per 1,000 residents),
with an estimated 14,000 deaths. Malaria accounts for 29\% of all
deaths, and 42\% of deaths among those under five years of age (INE
2011). Since 2013, Mozambique has seen a gradual increase in the
incidence of malaria (Moonasar et al. 2016). 100\% of the malaria in
Mozambique is of the Plasmodium falciparum species, with Anopheles
funestus, gambiae, and arabiensis as the primary mosquito vectors of the
disease (WHO 2015).

A significant sector of the economy in Mozambique is dominated by a full
large-scale foreing direct investment projects (Robbins and Perkins
2012), and the role of the private sector in health generally, and
malaria specifically, is unequivocally important. Large agriculture and
extractive industry firms take up wide swaths of land and employ
hundreds of thousands (German, Schoneveld, and Mwangi 2013). The
Mozambican state has encouraged large-scale entreprise with the aim of
general economic development (Buur, Tembe, and Baloi 2012). And where
large firms exist, they often take on social roles such as housing and
health care (Winkler 2013). At times, this role is necessary from a
purely practical standpoint; in other cases, it is employed under the
guise of ``corporate social responsibility'' (Azemar and Desbordes
2009). Regardless of the language used, it is clear that private
industry plays an important role in public health in Mozambique (Robbins
and Perkins 2012) (Castel-Branco 2014).

Several cases exist of foreign firms engaging in large-scale malaria
control campaigns (Mouzin and al. 2011) (Han 2015) (Bennett et al. 2017)
(Kaula, Buyungo, and Opigo 2017). But these studies generally consider
population health as the outcome measure of interest, rather than worker
absenteeism or productivity. Similarly, they often neglect to
differentiate between those clinical costs which are absorbed by the
local health system versus those which are absorbed by the firm itself.
In the literature, making the ``investment case'' for malaria control or
elimination generally implies that the investor is the public sector,
and takes into account those costs and benefits which are applicable
from a public welfare point of view (Shretta et al. 2017); though
appropriate in most cases (the government or institutions interested in
public welfare primarily being the primary malaria control agents in
most locations), the findings of these studies are rarely applicable to
the private sector. In the case of a private firm not interested in
``corporate social responsibility'', it is not clear whether investing
in malaria control would be profitable or not. This lack of clarity not
only discourages investment, but also makes it difficult for governments
to pinpoint the correct of amount of subsidy (if applicable) to
encourage private sector scale-up in malaria control.

@@

Our study adds to the existing literature in several ways.
\todo{Will add details here.}

@@

Maragra Açucar SA (henceforth referred to as
``Maragra'')\todo{This is a comment that will appear in the margin}

@@

In collaboration with the sugar processing facility, we collected data
for the period from January 2010 through December 2016. Data came from
four sources: (i) the Human Resources' roster of worker details and
absences, (ii) the facility's on-site clinic's medical and laboratory
records, (iii) the facility's on-site malaria control program's records
pertaining to the dates, chemicals, and location of IRS activities, and
(iv) interviews with company employees pertaining to costs, data
limitations, etc. Digitization and collection of data took place during
the period from March 2016 through May 2017. Supplementary data
pertaining to worker characteristics was obtained from through the
Centro de Investigação em Saude de Manhiça's (CISM) demographic census,
which covered workers from the district, but not those who migrated from
other parts of the country (Nhacolo et al. 2006).

@@

We sought to understand the effect of IRS on individual workers'
likelihood of absence from work as well as their likelihood of clinical
malaria. To estimate this effect, we estimated separate models for
absence and illness. We employed interrupted time series (Lopez Bernal,
Cummins, and Gasparrini 2016) and a linear probability approach using
the following econometric model.

@@

All data processing and analysis were carried out in R (R Core Team
2017) and all analysis code is freely available online (Brew 2017).
Ethical approval for this project was obtained from the Institutional
Ethics Review Board for Health at the CISM prior to data collection.

@@

\hypertarget{refs}{}
\hypertarget{ref-Azemar2009}{}
Azemar, C., and R. Desbordes. 2009. ``Public Governance, Health and
Foreign Direct Investment in Sub-Saharan Africa.'' \emph{Journal of
African Economies} 18 (4). Oxford University Press (OUP): 667--709.
doi:\href{https://doi.org/10.1093/jae/ejn028}{10.1093/jae/ejn028}.

\hypertarget{ref-Bennett_2017}{}
Bennett, Adam, Anton L. V. Avanceña, Jennifer Wegbreit, Chris Cotter,
Kathryn Roberts, and Roly Gosling. 2017. ``Engaging the Private Sector
in Malaria Surveillance: A Review of Strategies and Recommendations for
Elimination Settings.'' \emph{Malaria Journal} 16 (1). Springer Nature.
doi:\href{https://doi.org/10.1186/s12936-017-1901-1}{10.1186/s12936-017-1901-1}.

\hypertarget{ref-Bloom2008}{}
Bloom, David, and David Canning. 2008. ``Population Health and Economic
Growth,'' 1--25.

\hypertarget{ref-brewgit}{}
Brew, J.R. 2017. ``Malaria and Sugar: An in-Depth Examination of the
Effect of Malaria Control Activities on the Health and Productivity of
Maragra Sugarcane Factory Workers.'' \emph{GitHub Repository}.
\url{https://github.com/joebrew/maragra}; GitHub.

\hypertarget{ref-World1999}{}
Brundtland, Gro Harlem. 1999. ``WHO on Health and Economic
Productivity'' 25 (2): 396--402.

\hypertarget{ref-Buur2012}{}
Buur, Lars, Carlota Mondlane Tembe, and Obede Baloi. 2012. ``The White
Gold: The Role of Government and State in Rehabilitating the Sugar
Industry in Mozambique.'' \emph{Journal of Development Studies} 48 (3).
Informa UK Limited: 349--62.
doi:\href{https://doi.org/10.1080/00220388.2011.635200}{10.1080/00220388.2011.635200}.

\hypertarget{ref-CastelBranco2014}{}
Castel-Branco, Carlos Nuno. 2014. ``Growth, Capital Accumulation and
Economic Porosity in Mozambique: Social Losses, Private Gains.''
\emph{Review of African Political Economy} 41 (sup1). Informa UK
Limited: S26--S48.
doi:\href{https://doi.org/10.1080/03056244.2014.976363}{10.1080/03056244.2014.976363}.

\hypertarget{ref-German2013}{}
German, Laura, George Schoneveld, and Esther Mwangi. 2013.
``Contemporary Processes of Large-Scale Land Acquisition in Sub-Saharan
Africa: Legal Deficiency or Elite Capture of the Rule of Law?''
\emph{World Development} 48 (August). Elsevier BV: 1--18.
doi:\href{https://doi.org/10.1016/j.worlddev.2013.03.006}{10.1016/j.worlddev.2013.03.006}.

\hypertarget{ref-Goodman1999}{}
Goodman, CA, PG Coleman, and AJ Mills. 1999. ``Cost-Effectiveness of
Malaria Control in Sub-Saharan Africa.'' \emph{The Lancet} 354 (9176).
Elsevier BV: 378--85.
doi:\href{https://doi.org/10.1016/s0140-6736(99)02141-8}{10.1016/s0140-6736(99)02141-8}.

\hypertarget{ref-Han}{}
Han, Lily. 2015. ``Malaria in Mozambique: trialling payment by
results.''
\url{http://www.theguardian.com/global-development-professionals-network/2014/mar/31/malaria-control-payment-by-results}.

\hypertarget{ref-Hanson2004}{}
Hanson, K. 2004. ``Public and Private Roles in Malaria Control: The
Contributions of Economic Analysis.'' \emph{The American Journal of
Tropical Medicine and Hygiene} 71 (9176). The American Society of
Tropical Medicine; Hygiene: 168--73.
\url{http://www.ajtmh.org/docserver/fulltext/14761645/71/2_suppl/0700168.pdf?expires=1503070750\&id=id\&accname=guest\&checksum=F969ACD7BE7E3A5D929D87EEA8A6EA2B}.

\hypertarget{ref-Howard_2017}{}
Howard, Natasha, Lorna Guinness, Mark Rowland, Naeem Durrani, and
Kristian S. Hansen. 2017. ``Cost-Effectiveness of Adding Indoor Residual
Spraying to Case Management in Afghan Refugee Settlements in Northwest
Pakistan During a Prolonged Malaria Epidemic.'' Edited by Guilherme
S.Editor Ribeiro. \emph{PLOS Neglected Tropical Diseases} 11 (10).
Public Library of Science (PLoS): e0005935.
doi:\href{https://doi.org/10.1371/journal.pntd.0005935}{10.1371/journal.pntd.0005935}.

\hypertarget{ref-estatistica2009}{}
INE. 2011. ``Demographic Health Survey.'' \url{http://www.ine.gov.mz/}.

\hypertarget{ref-Kaula_2017}{}
Kaula, Henry, Peter Buyungo, and Jimmy Opigo. 2017. ``Private Sector
Role, Readiness and Performance for Malaria Case Management in Uganda,
2015.'' \emph{Malaria Journal} 16 (1). Springer Nature.
doi:\href{https://doi.org/10.1186/s12936-017-1824-x}{10.1186/s12936-017-1824-x}.

\hypertarget{ref-Lee2017}{}
Lee, Bruce Y., Eli Zenkov, Chandrani Chatterjee, Baltazar Candrinho,
Shufang Zhang, James Colborn, Olivier J. T. Briët, et al. 2017. ``The
Economic Value of Long-Lasting Insecticidal Nets and Indoor Residual
Spraying Implementation in Mozambique.'' \emph{The American Journal of
Tropical Medicine and Hygiene} 96 (6). American Society of Tropical
Medicine; Hygiene: 1430--40.
doi:\href{https://doi.org/10.4269/ajtmh.16-0744}{10.4269/ajtmh.16-0744}.

\hypertarget{ref-Lopez_Bernal_2016}{}
Lopez Bernal, James, Steven Cummins, and Antonio Gasparrini. 2016.
``Interrupted Time Series Regression for the Evaluation of Public Health
Interventions: A Tutorial.'' \emph{International Journal of
Epidemiology}, June. Oxford University Press (OUP), dyw098.
doi:\href{https://doi.org/10.1093/ije/dyw098}{10.1093/ije/dyw098}.

\hypertarget{ref-McCarthy_2000}{}
McCarthy, Desmond, Holger Wolf, and Yi Wu. 2000. ``The Growth Costs of
Malaria,'' February. National Bureau of Economic Research.
doi:\href{https://doi.org/10.3386/w7541}{10.3386/w7541}.

\hypertarget{ref-Moonasar_2016}{}
Moonasar, Devanand, Rajendra Maharaj, Simon Kunene, Baltazar Candrinho,
Francisco Saute, Nyasatu Ntshalintshali, and Natashia Morris. 2016.
``Towards Malaria Elimination in the Mosaswa (Mozambique, South Africa
and Swaziland) Region.'' \emph{Malaria Journal} 15 (1). Springer Nature.
doi:\href{https://doi.org/10.1186/s12936-016-1470-8}{10.1186/s12936-016-1470-8}.

\hypertarget{ref-Mouzin2011}{}
Mouzin, Eric, and Et al. 2011. ``Business Investing in Malaria Control:
Economic Returns and a Healthy Workforce for Africa.'' \emph{Progress \&
Impact Series}, no. 6.

\hypertarget{ref-Nhacolo_2006}{}
Nhacolo, Ariel Q, Delino A Nhalungo, Charfudin N Sacoor, John J Aponte,
Ricardo Thompson, and Pedro Alonso. 2006. ``Levels and Trends of
Demographic Indices in Southern Rural Mozambique: Evidence from
Demographic Surveillance in Manhiça District.'' \emph{BMC Public Health}
6 (1). Springer Nature.
doi:\href{https://doi.org/10.1186/1471-2458-6-291}{10.1186/1471-2458-6-291}.

\hypertarget{ref-Nonvignon_2016}{}
Nonvignon, Justice, Genevieve Cecilia Aryeetey, Keziah L. Malm, Samuel
Agyei Agyemang, Vivian N. A. Aubyn, Nana Yaw Peprah, Constance N.
Bart-Plange, and Moses Aikins. 2016. ``Economic Burden of Malaria on
Businesses in Ghana: A Case for Private Sector Investment in Malaria
Control.'' \emph{Malaria Journal} 15 (1). Springer Nature.
doi:\href{https://doi.org/10.1186/s12936-016-1506-0}{10.1186/s12936-016-1506-0}.

\hypertarget{ref-Orem_2012}{}
Orem, Juliet, Joses Kirigia, Robert Azairwe, Ibrahim Kasirye, and
Oladapo Walker. 2012. ``Impact of Malaria Morbidity on Gross Domestic
Product in Uganda.'' \emph{International Archives of Medicine} 5 (1).
Springer Nature: 12.
doi:\href{https://doi.org/10.1186/1755-7682-5-12}{10.1186/1755-7682-5-12}.

\hypertarget{ref-Purdy_2013}{}
Purdy, Mark, David Rublin, Kuangyi Wei, and Matthew Robinson. 2013.
``The Economic Case for Combating Malaria.'' \emph{The American Journal
of Tropical Medicine and Hygiene} 89 (5). American Society of Tropical
Medicine; Hygiene: 819--23.
doi:\href{https://doi.org/10.4269/ajtmh.12-0689}{10.4269/ajtmh.12-0689}.

\hypertarget{ref-R}{}
R Core Team. 2017. \emph{R: A Language and Environment for Statistical
Computing}. Vienna, Austria: R Foundation for Statistical Computing.
\url{https://www.R-project.org/}.

\hypertarget{ref-Robbins2012}{}
Robbins, Glen, and David Perkins. 2012. ``Mining Fdi and Infrastructure
Development on Africa's East Coast: Examining the Recent Experience of
Tanzania and Mozambique.'' \emph{Journal of International Development}
24 (2). Wiley-Blackwell: 220--36.
doi:\href{https://doi.org/10.1002/jid.2817}{10.1002/jid.2817}.

\hypertarget{ref-Sachs2002}{}
Sachs, Jeffrey, and Pia Malaney. 2002. ``The Economic and Social Burden
of Malaria.'' \emph{Nature} 415 (6872). Springer Nature: 680--85.
doi:\href{https://doi.org/10.1038/415680a}{10.1038/415680a}.

\hypertarget{ref-Shretta2016}{}
Shretta, Rima, Anton L. V. Avanceña, and Arian Hatefi. 2016. ``The
Economics of Malaria Control and Elimination: A Systematic Review.''
\emph{Malaria Journal} 15 (1). Springer Nature.
doi:\href{https://doi.org/10.1186/s12936-016-1635-5}{10.1186/s12936-016-1635-5}.

\hypertarget{ref-Shretta_2017}{}
Shretta, Rima, Ranju Baral, Anton L. V. Avanceña, Katie Fox, Asoka
Premasiri Dannoruwa, Ravindra Jayanetti, Arumainayagam Jeyakumaran,
Rasike Hasantha, Lalanthika Peris, and Risintha Premaratne. 2017. ``An
Investment Case to Prevent the Reintroduction of Malaria in Sri Lanka.''
\emph{The American Journal of Tropical Medicine and Hygiene}, January.
American Society of Tropical Medicine; Hygiene, 16--0209.
doi:\href{https://doi.org/10.4269/ajtmh.16-0209}{10.4269/ajtmh.16-0209}.

\hypertarget{ref-Vecchi_2013}{}
Vecchi, Veronica, Mark Hellowell, and Stefano Gatti. 2013. ``Does the
Private Sector Receive an Excessive Return from Investments in Health
Care Infrastructure Projects? Evidence from the Uk.'' \emph{Health
Policy} 110 (2-3). Elsevier BV: 243--70.
doi:\href{https://doi.org/10.1016/j.healthpol.2012.12.010}{10.1016/j.healthpol.2012.12.010}.

\hypertarget{ref-White_2011}{}
White, Michael T, Lesong Conteh, Richard Cibulskis, and Azra C Ghani.
2011. ``Costs and Cost-Effectiveness of Malaria Control Interventions -
a Systematic Review.'' \emph{Malaria Journal} 10 (1). Springer Nature:
337.
doi:\href{https://doi.org/10.1186/1475-2875-10-337}{10.1186/1475-2875-10-337}.

>>>>

\pagenumbering{gobble}
\usepackage{longtable}
\usepackage[utf8]{inputenc}
\usepackage{changepage}
\usepackage{graphicx}
\usepackage{multicol}
\usepackage{geometry}
\usepackage{fancyhdr}
\usepackage{color}
\usepackage{colortbl}
\usepackage{color}
% Font
\usepackage{fontspec}
\setmainfont{Swift-Regular_43151.ttf}
\setsansfont[BoldFont={Swift-Bold_43130.ttf}]{Swift-Regular_43151.ttf}
% \setmonofont{Swift-Regular_43151.ttf}
\renewcommand{\familydefault}{\sfdefault}
% \usepackage{fontspec}
% \setmainfont{Lato-Regular.ttf}
% \setsansfont[BoldFont={Lato-Bold.ttf}]{Lato-Regular.ttf}
% \renewcommand{\familydefault}{\sfdefault}

\def\changemargin#1#2{\list{}{\rightmargin#2\leftmargin#1}\item[]}
\let\endchangemargin=\endlist
\renewcommand{\rmdefault}{ppl}

\usepackage{multicol}
\usepackage{hyperref}
\usepackage{geometry}
\usepackage{lipsum}

% \usepackage{todonotes} % for side notes
% \usepackage[colorinlistoftodos]{todonotes} % for side notes

\usepackage{xargs}                      % Use more than one optional parameter in a new commands
\usepackage[dvipsnames, table]{xcolor}  % Coloured text etc.
% 
\usepackage[colorinlistoftodos,prependcaption,textsize=tiny]{todonotes}
\newcommandx{\unsure}[2][1=]{\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{#2}}
\newcommandx{\change}[2][1=]{\todo[linecolor=blue,backgroundcolor=blue!25,bordercolor=blue,#1]{#2}}
\newcommandx{\info}[2][1=]{\todo[linecolor=OliveGreen,backgroundcolor=OliveGreen!25,bordercolor=OliveGreen,#1]{#2}}
\newcommandx{\improvement}[2][1=]{\todo[linecolor=Plum,backgroundcolor=Plum!25,bordercolor=Plum,#1]{#2}}
\newcommandx{\thiswillnotshow}[2][1=]{\todo[disable,#1]{#2}}
\usepackage{lmodern}
\usepackage{fancyhdr} % Headers and footers
\pagestyle{fancy} % All pages have headers and footers
\fancyhead{} % Blank out the default header
\fancyfoot{} % Blank out the default footer
\fancyhead[C]{Return on investment of private sector malaria control at a large sugar facility in Southern Mozambique}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}

@@

\DeclareTextCommandDefault{\nobreakspace}{\leavevmode\nobreak\ }
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage[table]{xcolor}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}

@@

Malaria has a large economic impact on endemic societies. By affecting
saving, investment (Shretta et al., 2016), risk perception,
productivity, absenteeism (Nonvignon et al., 2016), human capital
accumulation (Castel-Branco, 2014), mortality, and costs of care (Sachs
and Malaney, 2002), malaria likely has a negative effect on GDP and
growth (McCarthy et al., 2000) (Orem et al., 2012). Because of the
relative affordability of most intereventions and the enormous societal
costs of malaria, most forms of malaria control are cost-effective when
a public welfare perspective is assumed, such as when a government
provides the financing (White et al., 2011) (Purdy et al., 2013) (Howard
et al., 2017).

From the persepctive of the private sector, however, investing in
malaria control is not so clear-cut. Public health interventions
targetting malaria - and their corresponding cost-effectiveness
evaluations - most often focus on impacts pertaining to public welfare,
such as an increase in life years adjusted for disability or quality
(Goodman et al., 1999) (Shretta et al., 2016) (Lee et al., 2017)
(Hanson, 2004). Though population-level health is certainly of
importance to businesses, and improvements in health incidentally
improve the economy at all levels (Brundtland, 1999) (Bloom and Canning,
2008) (Vecchi et al., 2013), these improvements may be too disperse or
long-term to incentivize private sector involvement in health campaigns.

100\% of the Mozambican population are at risk of malaria, living in
what the WHO classifies as a ``high transmission'' area (Moonasar et
al., 2016). Annually, Mozambique has more than 8 million clinical
malaria cases (an annual incidence of approximately 300 per 1,000
residents), with an estimated 14,000 deaths. Malaria accounts for 29\%
of all deaths, and 42\% of deaths among those under five years of age
(INE, 2011). Since 2013, Mozambique has seen a gradual increase in the
incidence of malaria (Moonasar et al., 2016). 100\% of the malaria in
Mozambique is of the Plasmodium falciparum species, with Anopheles
funestus, gambiae, and arabiensis as the primary mosquito vectors of the
disease (WHO, 2015).

A significant sector of the economy in Mozambique is dominated by a full
large-scale foreing direct investment projects (Robbins and Perkins,
2012), and the role of the private sector in health generally, and
malaria specifically, is unequivocally important. Large agriculture and
extractive industry firms take up wide swaths of land and employ
hundreds of thousands (German et al., 2013). The Mozambican state has
encouraged large-scale entreprise with the aim of general economic
development (Buur et al., 2012). And where large firms exist, they often
take on social roles such as housing and health care (Winkler, 2013). At
times, this role is necessary from a purely practical standpoint; in
other cases, it is employed under the guise of ``corporate social
responsibility'' (Azemar and Desbordes, 2009). Regardless of the
language used, it is clear that private industry plays an important role
in public health in Mozambique (Robbins and Perkins, 2012)
(Castel-Branco, 2014).

Several cases exist of foreign firms engaging in large-scale malaria
control campaigns (Mouzin and al., 2011) (Han, 2015) (Bennett et al.,
2017) (Kaula et al., 2017). But these studies generally consider
population health as the outcome measure of interest, rather than worker
absenteeism or productivity. Similarly, they often neglect to
differentiate between those clinical costs which are absorbed by the
local health system versus those which are absorbed by the firm itself.
In the literature, making the ``investment case'' for malaria control or
elimination generally implies that the investor is the public sector,
and takes into account those costs and benefits which are applicable
from a public welfare point of view (Shretta et al., 2017); though
appropriate in most cases (the government or institutions interested in
public welfare primarily being the primary malaria control agents in
most locations), the findings of these studies are rarely applicable to
the private sector. In the case of a private firm not interested in
``corporate social responsibility'', it is not clear whether investing
in malaria control would be profitable or not. This lack of clarity not
only discourages investment, but also makes it difficult for governments
to pinpoint the correct of amount of subsidy (if applicable) to
encourage private sector scale-up in malaria control.

@@

Our study adds to the existing literature in several ways.
\improvement{Will add details here.}

@@

Maragra Açucar SA (henceforth referred to as ``Maragra'').

@@

In collaboration with the sugar processing facility, we collected data
for the period from January 2010 through December 2016. Data came from
four sources: (i) the Human Resources' roster of worker details and
absences, (ii) the facility's on-site clinic's medical and laboratory
records, (iii) the facility's on-site malaria control program's records
pertaining to the dates, chemicals, and location of IRS activities, and
(iv) interviews with company employees pertaining to costs, data
limitations, etc. Digitization and collection of data took place during
the period from March 2016 through May 2017. Supplementary data
pertaining to worker characteristics was obtained from through the
Centro de Investigação em Saude de Manhiça's (CISM) demographic census,
which covered workers from the district, but not those who migrated from
other parts of the country (Nhacolo et al., 2006).

@@

We sought to understand the effect of IRS on individual workers'
likelihood of absence from work as well as their likelihood of clinical
malaria. To estimate this effect, we estimated separate models for
absence and illness. We employed interrupted time series (Lopez Bernal
et al., 2016) and a linear probability approach using the following
econometric model.

@@

All data processing and analysis were carried out in R (R Core Team,
2017) and all analysis code is freely available online (Brew, 2017).
Ethical approval for this project was obtained from the Institutional
Ethics Review Board for Health at the CISM prior to data collection.

@@

\hypertarget{refs}{}
\hypertarget{ref-Azemar2009}{}
Azemar, C., Desbordes, R., 2009. Public governance, health and foreign
direct investment in sub-saharan africa. Journal of African Economies
18, 667--709. \url{https://doi.org/10.1093/jae/ejn028}

\hypertarget{ref-Bennett_2017}{}
Bennett, A., Avanceña, A.L.V., Wegbreit, J., Cotter, C., Roberts, K.,
Gosling, R., 2017. Engaging the private sector in malaria surveillance:
A review of strategies and recommendations for elimination settings.
Malaria Journal 16. \url{https://doi.org/10.1186/s12936-017-1901-1}

\hypertarget{ref-Bloom2008}{}
Bloom, D., Canning, D., 2008. Population Health and Economic Growth
1--25.

\hypertarget{ref-brewgit}{}
Brew, J., 2017. Malaria and sugar: An in-depth examination of the effect
of malaria control activities on the health and productivity of maragra
sugarcane factory workers. GitHub repository.

\hypertarget{ref-World1999}{}
Brundtland, G.H., 1999. WHO on Health and Economic Productivity 25,
396--402.

\hypertarget{ref-Buur2012}{}
Buur, L., Tembe, C.M., Baloi, O., 2012. The white gold: The role of
government and state in rehabilitating the sugar industry in mozambique.
Journal of Development Studies 48, 349--362.
\url{https://doi.org/10.1080/00220388.2011.635200}

\hypertarget{ref-CastelBranco2014}{}
Castel-Branco, C.N., 2014. Growth, capital accumulation and economic
porosity in mozambique: Social losses, private gains. Review of African
Political Economy 41, S26--S48.
\url{https://doi.org/10.1080/03056244.2014.976363}

\hypertarget{ref-German2013}{}
German, L., Schoneveld, G., Mwangi, E., 2013. Contemporary processes of
large-scale land acquisition in sub-saharan africa: Legal deficiency or
elite capture of the rule of law? World Development 48, 1--18.
\url{https://doi.org/10.1016/j.worlddev.2013.03.006}

\hypertarget{ref-Goodman1999}{}
Goodman, C., Coleman, P., Mills, A., 1999. Cost-effectiveness of malaria
control in sub-saharan africa. The Lancet 354, 378--385.
\url{https://doi.org/10.1016/s0140-6736(99)02141-8}

\hypertarget{ref-Han}{}
Han, L., 2015. Malaria in Mozambique: trialling payment by results.

\hypertarget{ref-Hanson2004}{}
Hanson, K., 2004. Public and private roles in malaria control: The
contributions of economic analysis. The American Journal of Tropical
Medicine and Hygiene 71, 168--173.

\hypertarget{ref-Howard_2017}{}
Howard, N., Guinness, L., Rowland, M., Durrani, N., Hansen, K.S., 2017.
Cost-effectiveness of adding indoor residual spraying to case management
in afghan refugee settlements in northwest pakistan during a prolonged
malaria epidemic. PLOS Neglected Tropical Diseases 11, e0005935.
\url{https://doi.org/10.1371/journal.pntd.0005935}

\hypertarget{ref-estatistica2009}{}
INE, 2011. Demographic health survey.

\hypertarget{ref-Kaula_2017}{}
Kaula, H., Buyungo, P., Opigo, J., 2017. Private sector role, readiness
and performance for malaria case management in uganda, 2015. Malaria
Journal 16. \url{https://doi.org/10.1186/s12936-017-1824-x}

\hypertarget{ref-Lee2017}{}
Lee, B.Y., Zenkov, E., Chatterjee, C., Candrinho, B., Zhang, S.,
Colborn, J., Briët, O.J.T., Brown, S.T., Mendis, C., Bartsch, S.M.,
Viisainen, K., DePasse, J.V., Stone, N.T.B., 2017. The economic value of
long-lasting insecticidal nets and indoor residual spraying
implementation in mozambique. The American Journal of Tropical Medicine
and Hygiene 96, 1430--1440. \url{https://doi.org/10.4269/ajtmh.16-0744}

\hypertarget{ref-Lopez_Bernal_2016}{}
Lopez Bernal, J., Cummins, S., Gasparrini, A., 2016. Interrupted time
series regression for the evaluation of public health interventions: A
tutorial. International Journal of Epidemiology dyw098.
\url{https://doi.org/10.1093/ije/dyw098}

\hypertarget{ref-McCarthy_2000}{}
McCarthy, D., Wolf, H., Wu, Y., 2000. The growth costs of malaria.
\url{https://doi.org/10.3386/w7541}

\hypertarget{ref-Moonasar_2016}{}
Moonasar, D., Maharaj, R., Kunene, S., Candrinho, B., Saute, F.,
Ntshalintshali, N., Morris, N., 2016. Towards malaria elimination in the
mosaswa (mozambique, south africa and swaziland) region. Malaria Journal
15. \url{https://doi.org/10.1186/s12936-016-1470-8}

\hypertarget{ref-Mouzin2011}{}
Mouzin, E., al., E., 2011. Business Investing in Malaria Control:
Economic Returns and a Healthy Workforce for Africa. Progress \& Impact
series.

\hypertarget{ref-Nhacolo_2006}{}
Nhacolo, A.Q., Nhalungo, D.A., Sacoor, C.N., Aponte, J.J., Thompson, R.,
Alonso, P., 2006. Levels and trends of demographic indices in southern
rural mozambique: Evidence from demographic surveillance in manhiça
district. BMC Public Health 6.
\url{https://doi.org/10.1186/1471-2458-6-291}

\hypertarget{ref-Nonvignon_2016}{}
Nonvignon, J., Aryeetey, G.C., Malm, K.L., Agyemang, S.A., Aubyn,
V.N.A., Peprah, N.Y., Bart-Plange, C.N., Aikins, M., 2016. Economic
burden of malaria on businesses in ghana: A case for private sector
investment in malaria control. Malaria Journal 15.
\url{https://doi.org/10.1186/s12936-016-1506-0}

\hypertarget{ref-Orem_2012}{}
Orem, J., Kirigia, J., Azairwe, R., Kasirye, I., Walker, O., 2012.
Impact of malaria morbidity on gross domestic product in uganda.
International Archives of Medicine 5, 12.
\url{https://doi.org/10.1186/1755-7682-5-12}

\hypertarget{ref-Purdy_2013}{}
Purdy, M., Rublin, D., Wei, K., Robinson, M., 2013. The economic case
for combating malaria. The American Journal of Tropical Medicine and
Hygiene 89, 819--823. \url{https://doi.org/10.4269/ajtmh.12-0689}

\hypertarget{ref-R}{}
R Core Team, 2017. R: A language and environment for statistical
computing. R Foundation for Statistical Computing, Vienna, Austria.

\hypertarget{ref-Robbins2012}{}
Robbins, G., Perkins, D., 2012. Mining fdi and infrastructure
development on africa's east coast: Examining the recent experience of
tanzania and mozambique. Journal of International Development 24,
220--236. \url{https://doi.org/10.1002/jid.2817}

\hypertarget{ref-Sachs2002}{}
Sachs, J., Malaney, P., 2002. The economic and social burden of malaria.
Nature 415, 680--685. \url{https://doi.org/10.1038/415680a}

\hypertarget{ref-Shretta2016}{}
Shretta, R., Avanceña, A.L.V., Hatefi, A., 2016. The economics of
malaria control and elimination: A systematic review. Malaria Journal
15. \url{https://doi.org/10.1186/s12936-016-1635-5}

\hypertarget{ref-Shretta_2017}{}
Shretta, R., Baral, R., Avanceña, A.L.V., Fox, K., Dannoruwa, A.P.,
Jayanetti, R., Jeyakumaran, A., Hasantha, R., Peris, L., Premaratne, R.,
2017. An investment case to prevent the reintroduction of malaria in sri
lanka. The American Journal of Tropical Medicine and Hygiene 16--0209.
\url{https://doi.org/10.4269/ajtmh.16-0209}

\hypertarget{ref-Vecchi_2013}{}
Vecchi, V., Hellowell, M., Gatti, S., 2013. Does the private sector
receive an excessive return from investments in health care
infrastructure projects? Evidence from the uk. Health Policy 110,
243--270. \url{https://doi.org/10.1016/j.healthpol.2012.12.010}

\hypertarget{ref-White_2011}{}
White, M.T., Conteh, L., Cibulskis, R., Ghani, A.C., 2011. Costs and
cost-effectiveness of malaria control interventions - a systematic
review. Malaria Journal 10, 337.
\url{https://doi.org/10.1186/1475-2875-10-337}

\hypertarget{ref-whoprof}{}
WHO, 2015. Malaria profile: Mozambique.

\hypertarget{ref-Winkler}{}
Winkler, D., 2013. Potential and Actual FDI Spillovers in Global Value
Chains.
